---
title: "Kickstarter Project - CART Models Trainings"
output: html_notebook
---

This notebook briefly walks through creating a Kickstarter project success prediction model using CART.

First of all, we load all the necessary libraries.

```{r}
library(tidyverse)
library(rpart) # CART
library(rpart.plot) # CART plotting
library(caret) 
```

Then we load the train and validation data prepared in previous scripts.

```{r}
train <- readRDS('../rdata/df_initital_success_train.rds')
validate <- readRDS('../rdata/df_initital_success_validate.rds')
```


Using caret, a parameter sweep of `cp` was conducted to search for the value that provides that highest accuracy.

```{r}
if (file.exists("../model/cart.rds")) {
  cart_mod <- readRDS("../model/cart.rds")
} else {
  cpVals = data.frame(cp = seq(.00001, .0002, by=.00001))
  cart_mod <- train(as.factor(state)~., 
                      data = train,
                      method = "rpart",
                      tuneGrid = cpVals,
                      trControl = trainControl(method = "cv", number=10),
                      metric = "Accuracy")
  saveRDS(cart_mod, file="../model/cart.rds")
}
cart_mod$bestTune
```

A plot is created to see the accuracy of various models using different `cp`. A value of 0.00011 provides the highest accuracy.

```{r}
ggplot(cart_mod$results, aes(x=cp, y=Accuracy)) + geom_point(size=3) +
  xlab("Complexity Parameter (cp)") + geom_line()
```

With the best CART model, a classification tree was plotted. However the tree was way too complex to be interpretable. 

```{r}
#Best cart model 
best_cart_mod <- cart_mod$finalModel

#Classification Tree Plot
prp(best_cart_mod, digits=3)
```

We could also obtain the variable importance from the best CART model.

```{r}
# get variable importance as dataframe
impt <- best_cart_mod$variable.importance
var.impt <- as.data.frame(impt)
var.impt$importance <- var.impt$impt
var.impt$features <- names(best_cart_mod$variable.importance)

# plot importance
ggplot(var.impt, aes(x=features, y=importance)) + geom_bar(stat='identity') + coord_flip() + scale_x_discrete(limits=rev(var.impt$features))
```

As seen in the graph above, `unemployment rate` is the most important feature, with `tier.low` and `goal` as second and third.

Finally, let's see how does the model perform on validation data.

```{r}
#Cart Predictions with Best Tune Model
cart_pred <- predict(best_cart_mod, newdata=validate, type="class")
table(validate$state, cart_pred)

#CART Accuracy
mean(cart_pred == validate$state)

```