---
title: "Kickstarter Project - Random Forest Models Trainings"
output: html_notebook
---

This notebook briefly walks through creating a Kickstarter project success prediction model using random forest.

First of all, we load all the necessary libraries.

```{r}
library(caret) 
library(randomForest)
```

Then we load the train and validation data prepared in previous scripts.

```{r}
train <- readRDS('../rdata/df_initital_success_train.rds')
validate <- readRDS('../rdata/df_initital_success_validate.rds')
```

The train data is split into features (*xtrain*) and labels (*ytrain*).

```{r}
train$state <- as.factor(train$state)
xtrain <- train[,1:ncol(train)-1]
ytrain <- train$state
```

Using caret, a parameter sweep of `mtry` was conducted to search for the value that provides that highest accuracy.

```{r}
mtryValues <-  data.frame(mtry = seq(1, 20, by=1))
if(file.exists('../model/random_forest_train.rds')) {
  rf_train <- readRDS('../model/random_forest_train.rds')
} else {
  rf_train <- train(x = xtrain,
                 y = ytrain,
                 method = 'rf',
                 nodesize = 5,
                 ntree = 100,
                 tuneGrid = mtryValues,
                 trControl = trainControl(method = "cv", number=5),
                 metric = 'Accuracy',
                 do.trace=10
                 )
  saveRDS(rf_train, file='../model/random_forest_train.rds')
}
rf_train$bestTune
```

A plot is created to see the accuracy of various models using different `mtry`. A value of 6 provides the highest accuracy.

```{r}
ggplot(rf_train$results, aes(x=mtryValues, y=Accuracy)) + geom_point(size=3) +
  xlab("mtry") + geom_line()
```

Using the model with `mtry=6`, it was evaluated with the validation data.

```{r}
rf_mod <- rf_train$finalModel

rf_pred <- predict(rf_mod, newdata = validate)

table(validate$state, rf_pred)
print("Random Forest Accurracy")
mean(rf_pred == validate$state)
```

Accuracy is around 79 %. Another model is created with `mtry=6`, but with `ntree` increased to 500 and see whether that would improve the model.

```{r}
rf_node<-randomForest(
x = xtrain,
y = ytrain,
method = 'rf',
nodesize = 5,
ntree = 500,
mtry = 6,
metric = 'Accuracy',
do.trace = TRUE)
```

```{r}
#Playing around with the node size, above and below 5 always decreased accuracy
rf_node_pred <- predict(rf_node, newdata = validate)


#did not help
table(validate$state, rf_node_pred)
print("Random Forest Accurracy")
mean(rf_node_pred == validate$state)

```